Performance comparison of Dask and Apache Spark on HPC systems

### 2021-06-04 latex/paper.tex
[REWORK] methods section.

\title{Impact of engines scheduler on the performance of Big Data pipelines in HPC \TG{Suggested title: Performance comparison of Dask and Apache Spark on HPC systems}}

@@

\begin{IEEEkeywords}
	Performance, Scheduler, Big Data, Dask, Spark, Neuroimaging
\end{IEEEkeywords}
\MD{Discuss which keyword to use for the paper.}

@@

In the Dask Distributed scheduler, a \textit{dask-scheduler} process administrates the resource provided by  \textit{dask-worker}s in the cluster.
The scheduler receives jobs from clients and assigns tasks to available workers.
Task scheduler uses a FIFO (First-In-First-Out) job scheduling policy.
That is an utter process branch of the Dask graph before proceeding with the next one.

@@

Dask \MD{Add version when after locking it for experiments.} was used. \TG{should be in methods}

@@

In the Spark Standalone scheduler, a \textit{leader} \TG{I am all in favor of inclusive terminology, but if Spark didn't update their wording we shouldn't do it for them} coordinates the resource provisioned by \textit{workers} in the cluster.
A \textit{driver}  process receives jobs from clients and requests workers from the leader.
Jobs are divided into stages to be executed onto workers.
Each operation in a stage is represented by a high-level task in the computation graph.
Like Dask, Spark Standalone scheduler uses a FIFO policy to schedule tasks.
Spark Standalone has two execution modes: (1) the client mode, where the driver process runs in a dedicated process,
and (2) the cluster mode, where the driver runs within a worker process.
Our experiments use the client mode since cluster mode is not available in PySpark.

@@

\section{Methods}
\subsection{Infrastructure}
For our experiments we used the SlashBin cluster at Concordia University.
Each compute node has 2 16-cores CPUs,
\SI{256}{\giga\byte} memory, \MD{find out what's the memory speed and \#-channel supported.}
\SI{2.88}{\tera\byte} SSDs of mounted storage.
The nodes are interconnected with a dedicated \SI{10}{\giga\bit/\second} Ethernet connection.
Centos \MD{Verify version} and Linux kernel \MD{Verify kernel version} is installed on the nodes.

\MD{Write about the worker/leader setup once the dataset size is determined. Need to tune so that 2-3 blocks can fit in the excutors memmory}

\subsection{Dataset}
We used BigBrain\cite{Amunts:13}, a 3-D image of the human brain with voxel intensity ranging between 0 and 65,535.
We converted the blocks into the NIfTI format, a popular format in neuroimaging.
We left the NIfTI blocks uncompressed, resulting in a total data size of \SI{00}{\giga\byte}\MD{Verify the size of the dataset when Lustre back on}.
To evaluate the effect of block size, we resplit these blocks into 30, 125 and 750 blocks of \MD{size for each block segment}, using the \href{https://github.com/big-data-lab-team/sam}{sam} library.

We also used the dataset provided by the Consortium for Reliability and Reproducibility (\href{http://fcon_1000.projects.nitrc.org/indi/CoRR/html/}{CoRR}) \cite{zuo2014open}.
This dataset is freely available on \href{https://datasets.datalad.org/?dir=/corr/RawDataBIDS}{datalad}.
The entire dataset is \SI{408.4}{\giga\byte}, containing anatomical, diffusion and functional images of 1,397 subjects acquiredin 29 sites.
We used all 3,491 anatomical images, representing \SI{39}{\giga\byte} overall (\SI{11.17}{\mega\byte} per image on average).

\subsection{Applications}
\subsubsection{Increment}
We adapted the increment application used in \cite{hayot2019performance}.
This synthetic application reads blocks of the BigBrain from Lustre and simulates computation by sleeping for a specified period.
To simulate intermediate results, we repeat the sleep process for a configurable amount of time.
We prevent data caching of the blocks by incrementing their voxels value by one after each sleep operation.
Finally, we write the resulting NIfTI image back to Lustre.
This application allows us to study the engines when their inputs are processed independently.
The map-only scenario of this application mimics the processing of multiple independent subjects in parallel.
\MD{TODO: Add the Figure for the application graph.}

\subsubsection{Multi-Increment}
Our second application is an adaptation of the increment application.
A significant difference is that, at each iteration, it uses a random BigBrain block as the increment value. 
This change allows the multi-increment application to have inter-worker communication while remaining simple.
\MD{TODO: Add the Figure for the application graph.}

\subsubsection{Histogram}
As our third application, we calculate the histogram of the BigBrain image. 
The application reads the BigBrain blocks from Lustre, calculates each intensity's frequency, and then writes the aggregated result back on Lustre.
This map-reduce application has a very high read overwrite ratio.
Moreover, this application requires shuffling, albeit of a limited amoutn of data. 
The amount of inter-worker communication is in-between the increment and multi-increment applications.
\MD{TODO: Add the Figure for the application graph.}

\subsubsection{Kmeans} % TODO
For our fourth application, we apply Kmeans clustering to the voxel intensities of the BigBrain image.
We set the number of clusters to 3, to segment the white and grey matter and the noise.
The application starts by reading the image blocks, combining all voxels in a 1-D array, and choosing initial centroids using the min, max, and intermediate values.
It assigns each voxel to the centroids it is the closest and updates each centroid by computing the average of the voxels associated with it. 
It repeats the assignment and update steps for a configurable amount of time. 
Finally, the voxels of the image blocks are classified and written back to the file system.
Updating the centroids involves substantial data communication between the workers.

For this application, the Spark and Dask implementations differ slightly,  to take advantage of the best-suited API from both engines.
The Spark implementation uses the Map-Reduce paradigm, while the Dask one uses array programming.
\MD{TODO: Add the Figure for the application graph.}

\subsubsection{BIDS app example}
Our last application is BIDS App example: a neuroimaging pipeline to measure the brain volume from MRIs.
For this application, we use the CoRR dataset.
The application extracts the brain volume of each participant, then computes the average for each group of participants.
Unlike the other applications, BIDS App example is a command-line executed in a Docker image (bids/example on DockerHub).
We converted the Docker image to a Singularity image for use in HPC environments, \MD{Cite paper on reason why this is done.}
using \href{https://hub.docker.com/r/singularityware/docker2singularity/tags/}{docker2singularity}
\MD{Need to discuss if we preload the image on the worker nodes.}
\MD{TODO: Add the Figure for the application graph.}

\subsection{Experiments} % TODO

\section{Results} % TODO

\section{Discussion} % TODO

\section{Conclusion} % TODO

\section*{Acknowledgment} % TODO

\bibliographystyle{IEEEtran}
\bibliography{IEEEabrv,reference}

\end{document}

>>>>

\title{Performance comparison of Dask and Apache Spark on HPC systems}

@@

\begin{IEEEkeywords}
	Performance, Big Data, Dask, Spark, Neuroimaging
\end{IEEEkeywords}
\MD{Discuss which keyword to use for the paper.}

@@

In the Dask Distributed scheduler, a \textit{dask-scheduler} process administrates the resource provided by  \textit{dask-worker}s in the cluster.
The scheduler receives jobs from clients and assigns tasks to available workers.
Task scheduler uses a LIFO (Last-In-First-Out) job scheduling policy.
That is an utter process branch of the Dask graph before proceeding with the next one.

@@

@@

In the Spark Standalone scheduler, a \textit{leader} \TG{I am all in favor of inclusive terminology, but if Spark didn't update their wording we shouldn't do it for them} coordinates the resource provisioned by \textit{workers} in the cluster.
A \textit{driver}  process receives jobs from clients and requests workers from the leader.
Jobs are divided into stages to be executed onto workers.
Each operation in a stage is represented by a high-level task in the computation graph.
Like Dask, Spark Standalone scheduler uses a LIFO policy to schedule tasks.
Spark Standalone has two execution modes: (1) the client mode, where the driver process runs in a dedicated process,
and (2) the cluster mode, where the driver runs within a worker process.
Our experiments use the client mode since cluster mode is not available in PySpark.

@@

\section{Methods}
\subsection{Infrastructure}
For our experiments we used the SlashBin cluster at Concordia University.
Each compute node has 2 16-cores CPUs,
\SI{256}{\giga\byte} \SI{2666}{\mega\hertz} memory in 8-channel,
and \SI{2.88}{\tera\byte} SSDs of mounted storage.
The nodes are interconnected with a dedicated \SI{10}{\giga\bit/\second} Ethernet connection.
Centos 8 with Linux kernel \textit{4.18.0-240.1.1.el8\_lustre.x86\_64} installed on the compute nodes.

Both Spark and Dask are configured to have 8 worker processes each with 8 threads.
Each worker is allocated \SI{31.5}{\giga\byte} of memory.
A new cluster is spinned up and teared down for each experiment.

Dask ?? and Spark ?? \MD{Add version when after locking it for experiments.} was used for our experiments.

\subsection{Dataset}
We used BigBrain\cite{Amunts:13}, a 3-D image of the human brain with voxel intensity ranging between 0 and 65,535.
We converted the blocks into the NIfTI format, a popular format in neuroimaging.
We left the NIfTI blocks uncompressed, resulting in a total data size of \SI{648}{\giga\byte}.
To evaluate the effect of block size, we resplit these blocks into 1000, 2500, and 5000 blocks of \SI{648}{\mega\byte}, \SI{259.2}{\mega\byte}, and \SI{129.6}{\mega\byte}, respectively.
\href{https://github.com/big-data-lab-team/sam}{sam} library was used to resplit the image.

We also used the dataset provided by the Consortium for Reliability and Reproducibility (\href{http://fcon_1000.projects.nitrc.org/indi/CoRR/html/}{CoRR}) \cite{zuo2014open}, freely available on \href{https://datasets.datalad.org/?dir=/corr/RawDataBIDS}{datalad}.
The entire dataset is \SI{408.4}{\giga\byte}, containing anatomical, diffusion and functional images of 1,397 subjects acquiredin 29 sites.
We used all 3,491 anatomical images, representing \SI{39}{\giga\byte} overall (\SI{11.17}{\mega\byte} per image on average).

\subsection{Applications}
\subsubsection{Increment}
We adapted the increment application used in \cite{hayot2019performance}.
This synthetic application reads blocks of the BigBrain from Lustre and simulates computation by sleeping for a specified period.
To simulate intermediate results, we repeat the sleep process for a configurable amount of time.
We prevent data caching of the blocks by incrementing their voxels value by one after each sleep operation.
Finally, we write the resulting NIfTI image back to Lustre.
This application allows us to study the engines when their inputs are processed independently.
The map-only scenario of this application mimics the processing of multiple independent subjects in parallel.

\begin{figure}[!ht]
	\centering
	\includegraphics[height=\columnwidth,
	angle=0]{figures/increment.png}
	\caption{Task graph for Incrementation with 3 iterations and 3 BigBrain blocks.}
	\label{fig:graph-increment}
\end{figure}

% Currently ommitted due to issue with memory.
% 
\subsubsection{Multi-Increment}
Our second application is an adaptation of the increment application.
A significant difference is that, at each iteration, it uses a random BigBrain block as the increment value. 
This change allows the multi-increment application to have inter-worker communication while remaining simple.

\begin{figure}[!hb]
	\centering
	\includegraphics[height=\columnwidth,
	angle=0]{figures/multi-increment.png}
	\caption{Task graph for Multi-Incrementation with 3 iterations and 3 BigBrain blocks.}
	\label{fig:graph-muti-increment}
\end{figure}

\subsubsection{Histogram}
As our third application, we calculate the histogram of the BigBrain image. 
The application reads the BigBrain blocks from Lustre, calculates each intensity's frequency, and then writes the aggregated result back on Lustre.
This map-reduce application has a very high read overwrite ratio.
Moreover, this application requires shuffling, albeit of a limited amoutn of data. 
The amount of inter-worker communication is in-between the increment and multi-increment applications.

\begin{figure}[!hb]
	\centering
	\includegraphics[height=\columnwidth,
	angle=0]{figures/histogram.png}
	\caption{Task graph for Histogram with 3 BigBrain blocks.}
	\label{fig:graph-histogram}
\end{figure}

\subsubsection{Kmeans}
For our fourth application, we apply Kmeans clustering to the voxel intensities of the BigBrain image.
We set the number of clusters to 3, to segment the white and grey matter and the noise.
The application starts by reading the image blocks, combining all voxels in a 1-D array, and choosing initial centroids using the min, max, and intermediate values.
It assigns each voxel to the centroids it is the closest and updates each centroid by computing the average of the voxels associated with it. 
It repeats the assignment and update steps for a configurable amount of time. 
Finally, the voxels of the image blocks are classified and written back to the file system.
Updating the centroids involves substantial data communication between the workers.

For this application, the Spark and Dask implementations differ slightly,  to take advantage of the best-suited API from both engines.
The Spark implementation uses the Map-Reduce paradigm, while the Dask one uses array programming.

\subsubsection{BIDS App example}
Our fifth application is BIDS App example: a neuroimaging pipeline to measure the brain volume from MRIs.
For this application, we use the CoRR dataset.
The application extracts the brain volume of each participant, then computes the average for each group of participants.
Unlike the other applications, BIDS App example is a command-line executed in a Docker image (bids/example on DockerHub).
We converted the Docker image to a Singularity image for use in HPC environments, \MD{Cite paper on reason why this is done.}
using \href{https://hub.docker.com/r/singularityware/docker2singularity/tags/}{docker2singularity}

\subsubsection{BIDS App MRIQC}
Our last application is BIDS App MRIQC: a neuroimaging pipeline to perfom quality control of a brain image.
For this application, we use the CoRR dataset. \MD{Validate if this one was used after running the experiments. Maybe another dataset will be required; e.g. ADHD200}
This application verifies the quality of images on a per-sbuject basis.
Like the BIDS App example, this application is runned using a command-line tool.
We use the same method to convert the Docker image to a Singularity image.

\subsection{Experiments}
Table~\ref{table:parameters} the four parameters that are varied throughout the experiments.
We varied (1) the number of workers to assess the scalibility of the scheduler for the engine,
(2) the BigBrain block size in Increment, Multi-Increment, and Histogram to measure the effect of the different I/O pattern and parallelization degrees,
(3) the number of iterations to evaluate the effect of number of task,
and (4) the sleep delay to study the effect of task duration.
It should note that increasing the number of iterations for a given sleep delay also increases the total compute time of an application.

To avoid potential external bias such as caching, background process and network load, we ran the applications in randomized order and cleared the page cache in between every experiments.
Each benchmark was run ten times.

For each run, we measure the makespan of the application as well as the cumulative time spent in the different functions for read, processing, and writing data.
The overhead calculation for each CPU thread is the end time of the last processed task minus the total runtime of the tasks ran for this thread.
Summing those results gives the total overhead for the application.

\begin{table*}[t]
	\renewcommand{\arraystretch}{1.5}
	\caption{Parameters for the experiments}\label{table:parameters}
	\centering
	\begin{tabular}{|l|c|c|c|c|c|c|}
		\hline & Increment & Multi-Increment & Kmeans & Histogram & BIDS Example & BIDS MRIQC \\\hline
		\# of Nodes & \multicolumn{6}{c|}{2, 4, 8} \\\hline
		\# of Workers & \multicolumn{6}{c|}{16, 32 , 64} \\ \hline
		Block resplit & \multicolumn{4}{c|}{1000, 2500, 5000}  & \multicolumn{2}{c|}{n/a} \\\hline
		Block Size {[}\SI{}{\mega\byte}{]} & \multicolumn{4}{c|}{648, 259.2, 129.6} & \multicolumn{2}{c|}{n/a} \\\hline
		% \# of Subjects & \multicolumn{4}{c|}{n/a} & \multicolumn{2}{c|}{1397} \\ \hline
		\# of Iterations & \multicolumn{3}{c|}{1, 8, 64}                 & \multicolumn{3}{c|}{n/a} \\\hline
		Sleep Delay {[}\SI{}{\second}{]} & \multicolumn{2}{c|}{0.25, 1, 4, 16} & \multicolumn{4}{c|}{n/a} \\\hline
	\end{tabular}
\end{table*}
\MD{Decide which row to keep for (\# fo nodes / \# of workers) and (Block resplit / Block size)}

\section{Results} % TODO
\section{Discussion} % TODO
\section{Conclusion} % TODO
\section*{Acknowledgment} % TODO

\bibliographystyle{IEEEtran}
\bibliography{IEEEabrv,reference}
\end{document}
